---
title: "LA County Restaurant Inspection Scores"
author: "Victoria Yin"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
        code_folding: hide
---

LA County Restaurant Inspection Scores and Community Health

<br>

## Full Report [Here](https://github.com/v-yin/PM566-FinalProject/blob/main/FinalReport.pdf) [[Download]](https://github.com/v-yin/PM566-FinalProject/raw/main/FinalReport.pdf)

<br>

# Brief Introduction

Los Angeles County provides openly available data on all restaurant and market inspections over the past 5 years. Facilities are subject to inspection 1 to 3 times a year, and made public within 1 week of inspection date. The frequency in which restaurants and food markets are inspected depends on the public health risk associated with the food products served or prepared and on the facility's history of inspection grades. Inspectors deduct points based on violations and health risks, which is turned into a score out of 100. 


First, I examined all of the restaurants in the data set and calculated average inspection scores in each city to compare with its public health outcomes (Diabetes, Obesity, Depression). Then, I examined a subset of the restaurant inspection data (all McDonald's, Jack in the Box, Starbucks, El Pollo Loco, Panda Express, Taco Bell, Del Taco, In N Out, Panera Bread) and examined the geospatial correlation between restaurant inspection rating and its community health status.

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
# Initialize code chunk options
if(!require(data.table)) install.packages("data.table")
library(data.table)
if(!require(dplyr)) install.packages("dplyr")
library(dplyr)
if(!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)
if(!require(tidytext)) install.packages("tidytext")
library(tidytext)
if(!require(tidyverse)) install.packages("tidyverse")
library(tidyverse)
if(!require(dtplyr)) install.packages("dtplyr")
library(dtplyr)
if(!require(knitr)) install.packages("knitr")
library(knitr)
options(digits=2)
if(!require(forcats)) install.packages("forcats")
library(forcats)
if(!require(utils)) install.packages("utils")
library(utils)
if(!require(downloader)) install.packages("downloader")
library(downloader)
if(!require(ggmap)) install.packages("ggmap")
library(ggmap)
if(!require(leaflet)) install.packages("leaflet")
library(leaflet)
if(!require(leaflet.extras)) install.packages("leaflet.extras")
library(leaflet.extras)
if(!require(plotly)) install.packages("plotly")
library(plotly)
if(!require(DT)) install.packages("DT")
library(DT)
if(!require(magrittr)) install.packages("magrittr")
library(magrittr)
if(!require(RColorBrewer)) install.packages("RColorBrewer")
library(RColorBrewer)

opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")

# Read in and clean data
if (!file.exists("LACinspections.csv.zip"))
  download(
    url = "https://github.com/v-yin/PM566-FinalProject/raw/main/data/LACinspections.csv.zip",
    dest = "LACinspections.csv.zip",
    mode="wb"
    )
unzip("LACinspections.csv.zip", exdir="./")
inspect <- read.csv("LACinspections.csv")

if (!file.exists("LAChealth.csv"))
  download(
    url = "https://raw.githubusercontent.com/v-yin/PM566-FinalProject/main/data/LAChealth.csv",
    dest = "LAChealth.csv",
    mode="wb"
    )
health <- read.csv("LAChealth.csv")

health$GEONAME <- toupper(health$GEONAME)
health$GEONAME <- replace(health$GEONAME, health$GEONAME=="LOS ANGELES, CITY OF", "LOS ANGELES")

resthealth <- merge(
  x = inspect,
  y = health,
  by.x = "FACILITY_CITY",
  by.y = "GEONAME",
  all.x = FALSE,
  all.y = FALSE
)
resthealth <- data.table(resthealth)

# Delete scores less than 50
library(data.table)
resthealth <- resthealth[SCORE>50]
# Find restaurant chains
# resthealth %>% unnest_ngrams(word, FACILITY_NAME, n=2) %>% anti_join(stop_words, by = c("word")) # %>% count(word, sort=TRUE) %>% as_tibble() %>% print(n=100)
# Create chain variable
resthealth$FACILITY_NAME <- toupper(resthealth$FACILITY_NAME)
resthealth$CHAIN <- ifelse(grepl("MCDONALD", resthealth$FACILITY_NAME), "McDonald's", ifelse(grepl("JACK IN THE BOX", resthealth$FACILITY_NAME), "Jack in the Box", ifelse(grepl("STARBUCKS", resthealth$FACILITY_NAME), "Starbucks", ifelse(grepl("EL POLLO LOCO", resthealth$FACILITY_NAME), "El Pollo Loco", ifelse(grepl("PANDA EXPRESS", resthealth$FACILITY_NAME), "Panda Express", ifelse(grepl("TACO BELL", resthealth$FACILITY_NAME), "Taco Bell", ifelse(grepl("DEL TACO", resthealth$FACILITY_NAME), "Del Taco", ifelse(grepl("OUT BURGER", resthealth$FACILITY_NAME), "In N Out", ifelse(grepl("PANERA BREAD", resthealth$FACILITY_NAME), "Panera Bread", NA)))))))))
# Find average inspection score by chain
chain_avg <- resthealth[ , .(
  scoreavg = mean(SCORE) 
), by = "CHAIN"]
# Clean health outcome data to be numeric
resthealth$Prop_obse <- as.numeric(as.character(resthealth$Prop_obse))
resthealth$Prop_DM <- as.numeric(as.character(resthealth$Prop_DM))

library(tidygeocoder)

resthealth$fulladdress <- paste0(resthealth$FACILITY_ADDRESS, ", ", resthealth$FACILITY_CITY, ", CA ", resthealth$FACILITY_ZIP)

# restrict to chains only (too many data points to geocode)
# chains <- resthealth[!is.na(CHAIN)]
# chains <- as.data.frame(chains)

# chains <- chains %>% tidygeocoder::geocode(fulladdress, method = 'osm', lat = latitude , long = longitude)

# Take data set from github

if (!file.exists("chains.csv"))
  download(
    url = "https://github.com/v-yin/PM566-FinalProject/blob/main/data/chains.csv?raw=true",
    dest = "chains.csv",
    mode="wb"
    )
chains <- read.csv("chains.csv")
chains <- as.data.table(chains)
```

```{r initial data exploration, echo=FALSE, results = 'hide', message=FALSE}
library(ggplot2)
scorehisto <- ggplot(resthealth, aes(x=SCORE)) +
  geom_bar(fill="red") +
  ggtitle("Distribution of LA County Restaurant Inspection Scores, 2017-2022") +
  xlab("Inspection Score")


citysum <- resthealth %>% group_by(FACILITY_CITY) %>% summarise(mean_score = round(mean(SCORE), 2), sd_fev = round(sd(SCORE), 2), ninspect = n_distinct(RECORD_ID), Prop_DM, Prop_obse, Prop_depr) %>% distinct() 

citysum <- as.data.frame(citysum)
```

In total, there were `r nrow(resthealth)` inspections of `r n_distinct(resthealth$FACILITY_ID)` restaurants in `r n_distinct(resthealth$FACILITY_CITY)` cities within LA County.

Of the `r nrow(resthealth)` inspections included in the analysis, the average grade was `r mean(resthealth$SCORE)` with a standard deviation of `r sd(resthealth$SCORE)`. The highest score was a perfect score, `r max(resthealth$SCORE)` whereas the lowest score was `r min(resthealth$SCORE)`. 

```{r summarize}


summary <- as.data.frame(citysum)

datatable(summary, options = list(), class = "display",
    callback = JS("return table;"), rownames, colnames = c("City", "Average Score", "Standard Deviation of Score", "Number of Inspections in City", "Proportion Diabetes", "Proportion Obesity", "Proportion Depression"),
    caption = NULL, filter = c("none", "bottom", "top"), escape = TRUE,
    style = "auto", width = NULL, height = NULL, elementId = NULL,
    autoHideNavigation = getOption("DT.autoHideNavigation", NULL),
    selection = c("multiple", "single", "none"), extensions = list(),
    plugins = NULL, editable = FALSE)

```

Use the interactive table to explore how cities rank in terms of restaurant inspection scores and public health measures! The city with the highest inspection scores was Long Beach, followed by Calabasas. The city with the lowest inspection scores was Monterey Park.

<br>

```{r plotly}
fit1 <- lm(Prop_DM ~ mean_score, data = citysum)
scatter1 <- citysum %>% filter(!is.na(Prop_DM)) %>%
  plot_ly(x = ~mean_score, y = ~Prop_DM,
          type = 'scatter', mode = 'markers', marker = list(size = 5, color = colorRampPalette(brewer.pal(10,"Spectral"))(66))) %>%
          add_markers( hoverinfo = 'text',
          text = ~paste(paste(FACILITY_CITY),  paste0(" Average score:",
                        round(mean_score, 2), sep=""), paste(" Proportion DM:",
                        Prop_DM, sep=""))) %>%
  add_lines(x = ~mean_score, y=fitted(fit1)) %>%
  layout(title = "Average restaurant inspection score vs proportion of diabetes by city", yaxis = list(title = "Proportion Diabetes", hoverformat='.2f'), xaxis = list(title = "Average Inspection Score", hoverformat = '.2f'), hovermode = "compare") %>%
  layout(showlegend = F)

fit2 <- lm(Prop_obse ~ mean_score, data = citysum)
scatter2 <- citysum %>% filter(!is.na(Prop_obse)) %>%
  plot_ly(x = ~mean_score, y = ~Prop_obse,
          type = 'scatter', mode = 'markers', marker = list(size = 5, color = colorRampPalette(brewer.pal(10,"Paired"))(66))) %>%
          add_markers( hoverinfo = 'text',
          text = ~paste(paste(FACILITY_CITY),  paste0(" Average score:",
                        round(mean_score, 2), sep=""), paste(" Proportion Obese:",
                        Prop_obse, sep=""))) %>%
  add_lines(x = ~mean_score, y=fitted(fit2)) %>%
  layout(title = "Average restaurant inspection score vs proportion of obesity by city", yaxis = list(title = "Proportion Obesity"), xaxis = list(title = "Average Inspection Score"), hovermode = "compare") %>%
  layout(showlegend = F)

fit3 <- lm(Prop_depr ~ mean_score, data = citysum)
scatter3 <- citysum %>% filter(!is.na(Prop_depr)) %>%
  plot_ly(x = ~mean_score, y = ~Prop_depr,
          type = 'scatter', mode = 'markers', marker = list(size = 5, color = colorRampPalette(brewer.pal(10,"Pastel2"))(66))) %>%
          add_markers( hoverinfo = 'text',
          text = ~paste(paste(FACILITY_CITY),  paste0(" Average score:",
                        round(mean_score, 2), sep=""), paste(" Proportion with depression:",Prop_depr, sep=""))) %>%
  add_lines(x = ~mean_score, y=fitted(fit3)) %>%
  layout(title = "Average restaurant inspection score vs proportion with depression by city", yaxis = list(title = "Proportion Depression"), xaxis = list(title = "Average Inspection Score"), hovermode = "compare") %>%
  layout(showlegend = F)
```

## Showcasing scatterplots {.tabset}

### Restaurant Inspection Score vs Diabetes

```{r echo=FALSE, fig.show='asis'}
scatter1
```

There does not appear to be a significant linear correlation between a city's average restaurant inspection score and proportion of its population with diabetes.

### Restaurant Inspection Score vs Obesity
```{r echo=TRUE}
scatter2
```

There may be a weak positive correlation between a city's average inspection score and the proportion of its population with obesity.

### Restaurant Inspection Score vs Depression
```{r echo=TRUE}
scatter3
```

There appears to be a stronger positive correlation between a city's average inspection score and the proportion of its population with depression. Interestingly, Beverly Hills is a significant outlier here!

## {-}


